---
title: "condoR"
author: "Nicholas Knoblauch"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## How to use pajRo 

Let's say we have a really big csvfile, and it has a LOT of columns.  Too many columns!

```{r}
really_big_csvfile <- "../data-raw/lilteenyData.csv"
```

In this example file, we have bird sample data.  Each row has a `SAMPLING_EVENT_ID`. For every `SAMPLING_EVENT_ID` we have a bunch of metadata, and then we have a **column for like every kind of bird found anywhere in the dataset**. A `0` indicates that that bird was **not** found, and otherwise the number of birds found is indicated with something other than a `0` (it's usually a number, but sometimes it's an X).  What we'd rather have is a row for every non-zero bird sighting.

```{r}
library(pajRo)

regular_sized_data <- read_big_data(really_big_csvfile)
head(regular_sized_data)
```

So this is still a lot of columns.  Let's use `dplyr` (included in this package), and get some more info on which birds are found in which months

```{r}
library(dplyr)
month_data <- select(regular_sized_data,MONTH,Bird)
head(month_data)
```

Do we find more kinds of birds in certain months?

```{r}
group_by(month_data,MONTH) %>% summarise(n_birds=n_distinct(Bird))
```
Not really.
